---
title: "New York Time Mini Crossword"
author: "Stephen Kaluzny"
date: "August 11, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## The Puzzle

Everyday the New York Times puts out a mini crossword.

## Setup
```{r load_packages}
if(!requireNamespace("tidyverse", quietly=TRUE)) {
  install.packages("tidyverse")
}
library("tidyverse", quietly = TRUE, warn.conflicts = FALSE)
```
  
## The Data

```{r read_data}
d <- read.csv("nyt.csv", stringsAsFactors=FALSE)
```

Clean the data by dropping Player SKK, there is only one observation
at this time.

```{r clean}
nrow(d)
d <- d %>%
  filter(Player != "SKK")
nrow(d)
```
```{r date_time}
library("lubridate", quietly= TRUE)  # for working with date-time data
d$Date <- as.Date(d$Date)
d$WeekDay <- wday(d$Date)
d$Seconds <- with(d, as.numeric(seconds(ms(Time))))
```

## Summary Statistics

```{r summary}
d %>% group_by(Player) %>%
  summarise(Mean = mean(Seconds), Median = median(Seconds), Min = min(Seconds), Max = max(Seconds), N=n())
```

## Plots

An Initial plot of the data.

```{r plot01, echo=FALSE}
d %>%
  ggplot(aes(x=WeekDay, y=Seconds)) +
    geom_point() +
    ggtitle("Time vs Day of the Week")
```

Same plot with player identified.

```{r plot02, echo=FALSE}
d %>%
  ggplot(aes(x=WeekDay, y=Seconds, color=Player)) +
    geom_point() +
    ggtitle("Time vs Day of the Week")
```

The distribution of the times, summarised in a boxplot:

```{r plot03, echo=FALSE}
d %>%
  ggplot(aes(x=WeekDay, y=Seconds, group=WeekDay)) +
    geom_boxplot() +
    ggtitle("Time vs Day of the Week")