---
title: "Quality Check - NYT Mini Crossword Data"
author: "Stephen Kaluzny"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document: 
    theme: cerulean
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

We explicitly install the required R packages from CRAN if they
are not already available.

```{r load_packages}
library("assertr", quietly = TRUE, warn.conflicts = FALSE)
library("dplyr", quietly = TRUE, warn.conflicts = FALSE)
```
  
## The Data

The `here` package allows us to find files (in this case the raw data file)
anywhere within the project tree.

```{r read_data}
d <- read.csv(here::here("data", "nyt.csv"), stringsAsFactors=FALSE)
```

The `Date` and `TimeOfDay` are combined and converted to a `POSIX` variable `DateTime`.
The `WeekDay` is then computed from `DateTime`.
The puzzle time (`Time`) was recorded as minutes:seconds,
we convert that to a numeric `Seconds` variable.
```{r date_time}
d$Date <- as.Date(d$Date)
d$DateTime <- as.POSIXct(with(d,
  paste(Date, ifelse(is.na(TimeOfDay), "00:00", TimeOfDay))))
d$WeekDay <- lubridate::wday(d$DateTime)
d$TimeOfDayOrig <- d$TimeOfDay
d$TimeOfDay <- hms::parse_hm(d$TimeOfDay)
d$Seconds <- with(d, as.numeric(lubridate::seconds(lubridate::ms(Time))))
```

### Data Quality Check

New data is regularly added to the CSV data file.
In this section we do a quality check on the data to make sure no
data entry errors have occurred.

First, set some values that are expected in the data:

```{r data_verification_0}
# Current list of players:
players <- c("SPK", "JAK", "JIK", "BBK", "MBH", "SKK", "AKK", "MC", "DJK")
# Earliest date:
first_date <- as.Date("2017-09-25")
# Upper bound on time (10 minutes = 600 seconds):
time_bound <- 850
```

Check that Player is one of the `r length(players)` possible values:
```{r data_verification_1}
d %>% assert(in_set(players), Player, success_fun=success_logical,
  error_fun=error_logical)
```

There should be no missing values in the data:
```{r data_verification_2}
# d %>% assert(not_na, DateTime, WeekDay, Seconds) %>% success_logical()
d %>% assert(not_na, DateTime, WeekDay, Seconds,
  success_fun= success_logical, error_fun=error_logical)
```

Only one observation per player per date:
```{r data_verification_3}
# d %>% group_by(Date) %>% count(Player) %>% verify(n == 1) %>% success_logical()
d %>% group_by(Date) %>% count(Player) %>%
  verify(n == 1, success_fun=success_logical, error_fun=error_logical)
```

Check for proper time values:
```{r data_verification_4}
d %>% assertr::verify(Seconds > 0 & Seconds < time_bound,
  success_fun=success_logical, error_fun=error_logical)
```

Check range of dates
```{r data_verification_5}
first_date <- as.Date("2017-09-25")
today <- Sys.Date()
range(d$Date)
d %>% 
  assertr::verify(Date >= first_date & Date <= today,
    success_fun=success_logical, error_fun=error_logical)
```
